---
title: "Reproducible Research_Assignment1"
author: "Piyush Verma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## My First Assignement on R Markdown

This is my 1st assignment on R Markdown document.Here we will perform some actions given by the instructor.


### 1st task: Loading and preprocessing the data

Here we will be loading and processing the data, if necessary.



```{r Activity Monitoring}
library(dplyr)
library(ggplot2)
monitoring_activity<- read.csv("~/R files/activity.csv",header = TRUE)
monitoring_activity<- monitoring_activity[complete.cases(monitoring_activity),]
monitoring_activity$date<-as.Date(monitoring_activity$date,"%Y-%m-%d")

```

### 2nd task: Histogram of the total number of steps taken each day

Here we will plot the histogram of total number of steps taken each day with the help of plot function

```{r Total steps each day, echo=TRUE}
totalsteps<-tapply(monitoring_activity$steps,monitoring_activity$date,FUN = sum)
hist(totalsteps)
```

### 3rd task: Mean and median number of steps taken each day

Now we will calculate mean and median of number of steps each day.

```{r Mean and Median of steps each day, echo=TRUE}
meansteps<-mean(totalsteps)
mediansteps<-median(totalsteps)
cat("Mean number of steps: ", meansteps)
cat("Median number of steps: ", mediansteps)

```

### 4th task: Time series plot of the average number of steps taken

Now we will plot the daily activity pattern

```{r Time series plot of the average number of steps taken, echo=TRUE}

aggdata<- monitoring_activity %>% 
  group_by(interval) %>% 
  summarise(Avgsteps = mean(steps))
ggplot(aggdata,aes(x=interval,y=Avgsteps))+geom_point()

print("Max number of steps with the Interval")
aggdata[which.max(aggdata$Avgsteps),]

```

### 5th task: Inputing missing values with Avgsteps of that particular interval 

```{r Inputing missing values , echo=TRUE}

monitoring_activity2<- read.csv("~/R files/activity.csv",header = TRUE)
cat("total number of rows with NAs: ",sum(!complete.cases(monitoring_activity2)))


for(i in 1:nrow(monitoring_activity2))
    {

if (is.na(monitoring_activity2[i,1])==TRUE)
    {
    intervalNA<- monitoring_activity2[i,3]
    filterdata <- filter(aggdata,interval==intervalNA)
    monitoring_activity2[i,1]<-filterdata$Avgsteps
}
}


```

### 6th task: Doing all calculations 

```{r Doing all calculations again , echo=TRUE}

totalsteps2<-tapply(monitoring_activity2$steps,monitoring_activity2$date,FUN = sum)
hist(totalsteps2)
meansteps2<-mean(totalsteps2)
mediansteps2<-median(totalsteps2)
cat("Mean number of steps after imputting NA values: ", meansteps2)
cat("Median number of steps after imputting NA values: ", mediansteps2)

```

### 7th task: Classification of Date as Weekday/Weekend and Plotting of Avgsteps on Weekday vs Weekend

```{r Classification of Day as Weekday/Weekend , echo=TRUE}
monitoring_activity2$date<-as.Date(monitoring_activity2$date,"%Y-%m-%d")
monitoring_activity3<- mutate(monitoring_activity2, Day=weekdays(date))

Classification<-matrix(ncol = 1)
Classification<-data.frame(Classification)

for(i in 1:nrow(monitoring_activity3))
    {

if (monitoring_activity3[i,4]=="Saturday"|monitoring_activity3[i,4]=="Sunday")
    {
     Classification[i,] <- "Weekend"
}
    else{
         Classification[i,] <- "Weekday"
    }
}

monitoring_activity4<-mutate(monitoring_activity3,Classification)

aggdata2<- monitoring_activity4 %>% 
  group_by(Classification,interval) %>% 
  summarise(Avgsteps = mean(steps))
ggplot(aggdata2,aes(x=interval,y=Avgsteps,color=Classification))+geom_line()


```

